---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
suppressPackageStartupMessages({
  library(devout)
  library(ggplot2)
  library(dplyr)
  library(grid)
})
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "",
  fig.path = "man/figures/README-",
  out.width = "100%"
)


set.seed(1)
mtcars <- mtcars %>% mutate(cyl = as.factor(cyl))

```
# devout - non-standard devices for graphical output from R

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
![](https://img.shields.io/badge/Status-alpha-orange.svg)
![](https://img.shields.io/badge/Version-0.1.0-blue.svg)
<!-- badges: end -->


`devout` is a package for some of some non-standard devices for R.

A graphics device (e.g. `png()`, `pdf()`) is an interface to which R passes
plotting instructions. The graphics device then stores these in the appropriate
format for the given filetype.

However, there is no requirement that the output is saved to a file, nor is there
any requirement that the output is saved as a graphic.


## What's in the box

* `ascii()` - a non-standard device which outputs an ascii representation of the plot 
  to a file, or to the console or terminal
* `descriptive()` - a non-standard output device that describes what the output looks like

## Installation

You can install from [GitHub](https://github.com/coolbutuseless/devout) with:

``` r
# install.packages("remotes")
remotes::install_github("coolbutuseless/devout")
```


## `ascii()` device

The `ascii()` device draws an approximation of the graphics using ASCII characters
on the console (by default) or saved to a text file (if specified).

### Basic scatterplot

```{r}
p <- ggplot(mtcars) + 
  geom_point(aes(mpg, wt)) +
  labs(
    y     = "Car Weight",
    title    = "Basic scatter plot",
    subtitle = "Rendered with devout::ascii()"
  )  + 
  theme_bw()

ascii(width = 100)
p
jnk <- dev.off()
```

### Basic scatterplot with geom smooth

```{r}
p <- ggplot(mtcars) + 
  geom_point(aes(mpg, wt)) +
  geom_smooth(aes(mpg, wt), se = FALSE, method = 'loess') +
  labs(
    y        = "Car Weight",
    title    = "Basic scatter plot with a geom_smooth()",
    subtitle = "Rendered with devout::ascii()"
  )  + 
  theme_bw()

ascii(width = 100)
p
jnk <- dev.off()
```


### Boxplot with `coord_flip()`

```{r}
p <- ggplot(mtcars) + 
  geom_boxplot(aes(cyl, wt)) +
  labs(title = "Boxplot with flipped coorindates", 
    subtitle = "Rendered with devout::ascii()") + 
  coord_flip() + 
  theme_bw() 

ascii(width = 100, height = 30)
grid.draw(p)
jnk <- dev.off()
```


### Facetted Histogram with a `fill` aesthetic

```{r eval = TRUE}
p <- ggplot(mtcars) + 
  theme_bw()  +
  geom_histogram(aes(wt, fill = cyl), bins = 5, colour = 'black') +
  labs(title = "Facetted Histogram with 'fill' Aesthetic", 
    subtitle = "Rendered with devout::ascii()") + 
  facet_wrap(~cyl, labeller = label_both) +
  theme(panel.grid = element_blank())

ascii(width = 100)
grid.draw(p)
jnk <- dev.off()
```



## `descriptive()` device

The descriptive device gives a blow-by-blow account of what the graphics
device is being asked to draw.

This device is useful to use as a starting point for creating other graphics devices - 
I used it as the basis for the `ascii()` device and added featues as I needed them.



```{r}
p <- ggplot(mtcars) + 
  geom_point(aes(mpg, wt)) +
  labs(title = "Rendered with devout::ascii() more text goes here") +
  theme_void()

descriptive()
p
jnk <- dev.off()
```

## Limitations

* No filled polygons.
* No anti-aliasing.
* No UTF text support.
* No plotmath support.
* No sanity checking or safety checking of 'filename' string.
* No way to have un-filled circles.
* No stroke aesthetic for cicles.
* No real support for legends.
    * Legends get drawn, but a colourscale (for example) doesn't really map 
      to an ascii representation so it looks rubbish.
    * You should probably always add `theme(legend.position = 'none')` .
* No support for carriage returns in text.


## Other Non-Standard Device Ideas

* ANSI graphics
* Colour ASCII/ANSI output
* audio + midi output
* HPGL plotting output
* [CNC](https://en.wikipedia.org/wiki/Numerical_control) machine tooling instructions
* Directly drive motors to power an etch-a-sketch
* Crochet and knitting patterns
* Capture each call and turn it back into R code as a sequence of `grid` calls
* Save drawing instructions as a CSV file



## References:

* [svglite source code](https://github.com/r-lib/svglite)
* The defunct `RGraphicsDevice` project
    * http://www.omegahat.net/RGraphicsDevice/
    * http://www.omegahat.net/RGraphicsDevice/overview.html

