---
title: "Creating an SVG device - Part 3"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating an SVG device - Part 3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(xml2)
  library(devout)
  library(glue)
  library(ggplot2)
})
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


Introduction
==============================================================================

This series of vignettes will walk through the process of creating a very basic SVG
output device using `devout`.

The only code to be written is R code.  All the tricky C/C++ code is taken care
of within `devout`.

Vignettes in this series:

1. A simple callback function
2. Setting up a 'canvas' upon which to write the SVG
3. This vignette:
    * Adding support for device calls which draw on the canvas
4. Simple experimentation



Adding support for device calls which draw on the canvas
==============================================================================

In the previous vignettes we create a callback function (`svg_callback()`) and
implemented handlers for `device_call = "open"` and `"close"`.

Now we will add handlers for the device calls which want to draw a graphics 
primitive e.g. `rect`, `line`, `circle`.

The `svg_open()` and `svg_close()` functions are adapted to only write the lead-in
and lead-out parts of an SVG document.  The size for the SVG canvas is found 
in the device settings i.e. `state$dd`.  For a list of all the possible device settings, 
please see `devinfo$dd`.

<details closed>

<summary> <span title='dd info'> device settings (click to open) </summary> 

```{r}
devinfo$dd %>%
  knitr::kable(caption = "Device settings")
```

</details><br />


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# When opening a device
#  - create a "canvas".  For svg, the canvas is just a text string of SVG 
#    commands that we'll keep adding to with each device call
#  - add the canvas to the 'state$rdata' list
#  - always return the state so we keep the canvas across different device calls
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_open <- function(args, state) {
  state$rdata$svg <- glue::glue('<svg height="{state$dd$bottom/72}" width="{state$dd$right/72}">')
  state
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# When the device is closed, add the closing svg tag and output to file
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_close <- function(args, state) {  
  state$rdata$svg <- paste(state$rdata$svg, "</svg>", sep = "\n")
  writeLines(state$rdata$svg, state$rdata$filename)
  state
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Add a circle to the SVG
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_circle <- function(args, state) {
  state$rdata$svg <- paste(
    state$rdata$svg,
    glue::glue('<circle cx="{args$x/72}" cy="{args$y/72}" r="{args$r}" stroke="black" fill="black" />'),
    sep = "\n"
  )
  state
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Add a line to the SVG
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_polyline <- function(args, state) {
  state$rdata$svg <- paste(
    state$rdata$svg,
    glue::glue('<polyline points="{paste(args$x/72, args$y/72, sep=",", collapse = " ")}" stroke="black"  fill = "none" />'),
    sep = "\n"
  )
  state
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Add a polyline to the SVG
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_line <- function(args, state) {
  state$rdata$svg <- paste(
    state$rdata$svg,
    glue::glue('<line x1="{args$x1/72}" y1="{args$y1/72}" x2="{args$x2/72}" y2="{args$y2/72}" stroke="black"  fill = "none" />'),
    sep = "\n"
  )
  state
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Add text to the SVG
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_text <- function(args, state) {
  state$rdata$svg <- paste(
    state$rdata$svg,
    glue::glue('<text x="{args$x/72}" y="{args$y/72}" transform="rotate({-args$rot}, {args$x/72}, {args$y/72})" fill="black">{args$str}</text>'),
    sep = "\n"
  )
  state
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Call a different function to match each of the device calls we handle.
# Always return the state
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_callback <- function(device_call, args, state) {
  switch(
    device_call,
    "open"         = svg_open      (args, state),
    "close"        = svg_close     (args, state),
    "circle"       = svg_circle    (args, state),
    "line"         = svg_line      (args, state),
    "polyline"     = svg_polyline  (args, state),
    "textUTF8"     = svg_text      (args, state),
    state
  )
}
```


## Example plot 

```{r}
rdevice("svg_callback", filename = "svg/test-draw.svg") 
ggplot(mtcars) + geom_point(aes(mpg, wt))
invisible(dev.off())
```



View the SVG text output

```{r}
cat(paste(readLines("svg/test-draw.svg"), collapse = "\n"))
```

Open the output in an SVG viewer

```{r}
txt <- readLines("svg/test-draw.svg")
htmltools::HTML(txt)
```



# Next steps

This plot output doesn't look that great because we cut a lot of corners to get here; this includes:

* no clipping support
* no support for different colours for lines, points or filling. If you review
  the code you'll see that all line strokes are set to black, and only
  circles are filled.
* since `strWidth` and `metricInfo` aren't supported yet, the drawing system
  really doesn't know how to position the fonts well, or how to leave space for them
  near other objects.
  








