---
title: "Creating an SVG device - Part 4"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating an SVG device - Part 4}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(xml2)
  library(devout)
  library(glue)
  library(ggplot2)
})
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


Introduction
==============================================================================

This series of vignettes will walk through the process of creating a very basic SVG
output device using `devout`.

The only code to be written is R code.  All the tricky C/C++ code is taken care
of within `devout`.

Vignettes in this series:

1. A simple callback function
2. Setting up a 'canvas' upon which to write the SVG
3. Adding support for device calls which draw on the canvas
4. This vignette:
    * Simple experimentation
    * Just because we're told to draw a line doesn't mean we have to


%#$^ you I won't do what you tell me.
==============================================================================

In the previous vignettes we filled out the `svg_callback()` function with 
more support for drawing graphics primitives.

Now comes the interesting part: just because the device call wants us to draw a
line between two points, we don't have to do exactly as asked.  For example, instead
of drawing a single line from A to B, we could:

* Draw a wavy line
* Draw multiple jittered lines
* Draw a curve
* Replace the line with a sequence of circles
* Play `beepr::beep()` instead of drawing anything at all.
* Draw a spline
* Draw a tapered polygon to mimic a thick brush stroke

The following version of the graphics device adds multiple jittered lines for 
every line requested.  It doesn't look pretty, or very interesting, but hopefully 
shows that rule breaking is possible now that we control the graphics device!


```{r echo = FALSE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# When opening a device
#  - create a "canvas".  For svg, the canvas is just a text string of SVG 
#    commands that we'll keep adding to with each device call
#  - add the canvas to the 'state$rdata' list
#  - always return the state so we keep the canvas across different device calls
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_open <- function(args, state) {
  state$rdata$svg <- glue::glue('<svg height="{state$dd$bottom/72}" width="{state$dd$right/72}">')
  state
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# When the device is closed, add the closing svg tag and output to file
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_close <- function(args, state) {  
  state$rdata$svg <- paste(state$rdata$svg, "</svg>", sep = "\n")
  writeLines(state$rdata$svg, state$rdata$filename)
  state
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Add a circle to the SVG
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_circle <- function(args, state) {
  state$rdata$svg <- paste(
    state$rdata$svg,
    glue::glue('<circle cx="{args$x/72}" cy="{args$y/72}" r="{args$r}" stroke="black" fill="black" />'),
    sep = "\n"
  )
  state
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Add a line to the SVG
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_polyline <- function(args, state) {
  state$rdata$svg <- paste(
    state$rdata$svg,
    glue::glue('<polyline points="{paste(args$x/72, args$y/72, sep=",", collapse = " ")}" stroke="black"  fill = "none" />'),
    sep = "\n"
  )
  state
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Add text to the SVG
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_text <- function(args, state) {
  state$rdata$svg <- paste(
    state$rdata$svg,
    glue::glue('<text x="{args$x/72}" y="{args$y/72}" transform="rotate({-args$rot}, {args$x/72}, {args$y/72})" fill="black">{args$str}</text>'),
    sep = "\n"
  )
  state
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Add a polyline to the SVG
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_line <- function(args, state) {
  state$rdata$svg <- paste(
    state$rdata$svg,
    glue::glue('<line x1="{args$x1/72}" y1="{args$y1/72}" x2="{args$x2/72}" y2="{args$y2/72}" stroke="black"  fill = "none" />'),
    sep = "\n"
  )
  state
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Call a different function to match each of the device calls we handle.
# Always return the state
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_callback <- function(device_call, args, state) {
  switch(
    device_call,
    "open"         = svg_open      (args, state),
    "close"        = svg_close     (args, state),
    "circle"       = svg_circle    (args, state),
    "line"         = svg_line      (args, state),
    "polyline"     = svg_polyline  (args, state),
    "textUTF8"     = svg_text      (args, state),
    state
  )
}
```

```{r include=FALSE}
set.seed(1)
```


```{r}
svg_polyline <- function(args, state) {
  n <- length(args$x)
  state$rdata$svg <- paste(
    state$rdata$svg,
    glue::glue('<polyline points="{paste(args$x/72 + rnorm(n) * 3, args$y/72 + rnorm(n) * 3, sep=",", collapse = " ")}" stroke="black"  stroke-opacity="0.2" fill = "none" />'),
    glue::glue('<polyline points="{paste(args$x/72 + rnorm(n) * 3, args$y/72 + rnorm(n) * 3, sep=",", collapse = " ")}" stroke="black"  stroke-opacity="0.2" fill = "none" />'),
    glue::glue('<polyline points="{paste(args$x/72 + rnorm(n) * 3, args$y/72 + rnorm(n) * 3, sep=",", collapse = " ")}" stroke="black"  stroke-opacity="0.2" fill = "none" />'),
    sep = "\n"
  )
  state
}

```




## Example plot 

```{r}
rdevice(svg_callback, filename = "svg/test-experiment.svg", width=10*72, height = 6*72) 
ggplot(mtcars) + geom_point(aes(mpg, wt))
invisible(dev.off())
```




<details closed>
<summary> View the raw SVG text </summary>

```{r}
cat(paste(readLines("svg/test-experiment.svg"), collapse = "\n"))
```

</details> <br />


Open the output in an SVG viewer

```{r out.width="100%", fig.width=10}
txt <- readLines("svg/test-experiment.svg")
htmltools::HTML(txt)
```

<br />








