---
title: "Creating an SVG device - Part 2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating an SVG device - Part 2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(xml2)
  library(devout)
  library(glue)
  library(ggplot2)
})

knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  dev      = "svglite",
  fig.ext  = ".svg"
)
```


Introduction
==============================================================================

This series of vignettes will walk through the process of creating a very basic SVG
output device using `devout`.

The only code to be written is R code.  All the tricky C/C++ code is taken care
of within `devout`.

Vignettes in this series:

1. A simple callback function
2. This Vignette:
    * What is SVG?
    * Setting up a 'canvas' upon which to write the SVG
3. Adding support for device calls which draw on the canvas
4. Simple experimentation


What is SVG?
==============================================================================

"Scalable Vector Graphics (SVG) is an XML-based vector image format for two-dimensional graphics with support for interactivity and animation" - [wikipedia](https://en.wikipedia.org/wiki/Scalable_Vector_Graphics)

An SVG file is just a text file with a specific format which contains instructions
for drawing vector shapes e.g.

```
<?xml version="1.0" encoding="UTF-8" ?>
<svg width="391" height="391" viewBox="-70.5 -70.5 391 391" xmlns="http://www.w3.org/2000/svg">
  <rect x="25" y="25" width="200" height="200" fill="lime" stroke-width="4" stroke="pink" />
  <circle cx="125" cy="125" r="75" fill="black" />
  <polyline points="50,150 50,200 200,200 200,100" stroke="red" stroke-width="4" fill="none" />
  <line x1="50" y1="50" x2="200" y2="200" stroke="blue" stroke-width="4" />
</svg>
```



Set up a "canvas"
==============================================================================

Since SVG is a text format, our canvas for drawing instructions is just a
character string.

Out *SVG callback* from the prior vignette needs to be expanded so that:

* When `device_call = 'open'` is called, a new string is created.
* When `device_call = 'close'` is called, the string is written to file 

In this (and future examples), the `svg_callback` is a function which really
just contains a giant switch statement which then calls particular functions
for each `device_call`

```{r echo = TRUE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Call a different function to match each of the device calls we handle.
# Always return the state
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_callback <- function(device_call, args, state) {
  switch(
    device_call,
    "open"  = svg_open (args, state),
    "close" = svg_close(args, state),
    state
  )
}
```


In the following, `svg_open()` is called when `device_call = 'open'` and it 
creates a character string (called "svg") which contains a minimal SVG document.

Note that the string is created as part of the `state$rdata`.  It needs to be part
of the `state` as this is the only data structure which is passed to-and-from the R internals
and every device call.

```{r echo = TRUE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# When opening a device
#  - create a "canvas".  For svg, the canvas is just a text string of SVG 
#    commands that we'll keep adding to with each device call
#  - add the canvas to the 'state$rdata' list
#  - always return the state so we keep the canvas across different device calls
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_open <- function(args, state) {
  state$rdata$svg <- '
<svg height="100" width="100">
  <circle cx="50" cy="50" r="40" stroke="black" stroke-width="3" fill="red" />
</svg>'
  
  state
}
```


`svg_close()` is called when `device_call = 'close'` and this function
writes the canvas (`state$rdata$svg`) to a file - `state$rdata$filename`.


```{r echo = TRUE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# When the device is closed, output the SVG somehow
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
svg_close <- function(args, state) {
  writeLines(state$rdata$svg, state$rdata$filename)
  state
}
```


## Example plot

The filename to write to comes from the rdevice call.  The first argument is
always the name of the callback function.  Any further named arguments become
named elements of `state$rdata` and made available to the callback function.


```{r echo = TRUE}
rdevice(svg_callback, filename = "svg/test-canvas.svg") 
plot(1:10)
invisible(dev.off())
```


View the SVG text output

```{r}
cat(paste(readLines("svg/test-canvas.svg"), collapse = "\n"))
```

Open the output in an SVG viewer

```{r}
txt <- readLines("svg/test-canvas.svg")
htmltools::HTML(txt)
```




